---
title: "Deploying hySpc.hpc on a High Performance Computing (HPC) Cluster"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Deploying hySpc.hpc on a High Performance Computing (HPC) Cluster}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `hySpc.hpc` package derives its "High-Performance" name from its ability to dramatically accelerate spatial smoothing operations on hyperspectral images. It achieves this by bridging R with Rust, utilizing the `faer` crate for highly optimized sparse linear algebra, and the `rayon` crate for multi-threading.

However, `hySpc.hpc` is designed for **intra-node acceleration**. This means it utilizes all available CPU cores *on a single machine* to parallelize operations across the wavelength bands of a single hyperspectral image.

In many real-world scenarios—such as processing hundreds or thousands of distinct hyperspectral images—you will need **inter-node acceleration** (i.e., cluster computing). This vignette explains how to bridge the gap between `hySpc.hpc`'s internal multi-threading and an external job scheduler like **Slurm**, **SGE**, or **Torque** using the `future` ecosystem.

## The Architecture of Parallelism

When scaling hyperspectral processing, you are dealing with two levels of parallelism:

1. **Inner Loop (Rust/Rayon):** `hySpc.hpc` automatically distributes the smoothing of $k$ distinct wavelength bands across the CPU cores available on the machine executing the function.
2. **Outer Loop (Slurm/Future):** You want to distribute $N$ distinct hyperspectral images across multiple compute nodes in a cluster.

> **CRITICAL:** If you do not manage these two layers carefully, you will encounter the **Nested Parallelism Trap**. If a Slurm node is allocated 4 CPU cores for an R worker, but Rust's Rayon detects that the physical hardware has 64 cores, Rayon will spawn 64 threads. If you have multiple R workers on that same node, you will massively oversubscribe the CPU, causing severe performance degradation or job termination by the cluster administrator.

## Setting up `future.batchtools`

The most robust way to run R code on an HPC cluster is using the `future` framework alongside `future.batchtools`. This abstracts away the specifics of the job scheduler.

First, ensure you have the necessary packages:

```{r eval=FALSE}
install.packages(c("future", "future.batchtools", "future.apply"))
```

### Creating a Slurm Template

You need a template file (`slurm.tmpl`) that tells `batchtools` how to format the bash script sent to the `sbatch` command. Save this file in your working directory.

```bash
#!/bin/bash
#SBATCH --job-name=<%= job.name %>
#SBATCH --output=<%= log.file %>
#SBATCH --error=<%= log.file %>
#SBATCH --ntasks=1
#SBATCH --cpus-per-task=<%= resources$ncpus %>
#SBATCH --mem-per-cpu=<%= resources$memory %>
#SBATCH --time=<%= resources$walltime %>

Rscript -e 'batchtools::doJobCollection("<%= uri %>")'
```

### Configuring the R Environment

Next, configure the `future` plan in your R script to use this template.

```{r eval=FALSE}
library(future.batchtools)
library(future.apply)
library(hySpc.hpc)

# Define the compute resources for EACH worker job
# e.g., We want each worker to get 4 CPU cores and 8GB RAM per core.
resources <- list(
  ncpus = 4, 
  memory = 8000, 
  walltime = "02:00:00"
)

# Tell the future framework to use Slurm for asynchronous jobs
plan(batchtools_slurm, template = "slurm.tmpl", resources = resources)
```

## Dispatching `hySpc.hpc` to the Cluster

Assume we have a list of `hyperSpec` objects representing different biological samples or scans, and we want to apply spatial smoothing to all of them.

Here is the complete workflow to dispatch this safely to the cluster:

```{r eval=FALSE}
# Assume `spc_list` is a list of 100 hyperSpec objects
spc_list <- load_my_hyperspectral_data() 

# Define the smoothing function that will run on the cluster nodes
smooth_wrapper <- function(spc) {
  # -------------------------------------------------------------
  # CRITICAL: THREAD MANAGEMENT
  # We MUST tell Rust/Rayon to restrict its thread pool to the 
  # number of cores officially allocated to this job by Slurm.
  # -------------------------------------------------------------
  allocated_cores <- as.numeric(Sys.getenv("SLURM_CPUS_PER_TASK", unset = 1))
  Sys.setenv(RAYON_NUM_THREADS = allocated_cores)
  
  # Now it is safe to run the high-performance Rust backend
  # It will use exactly `allocated_cores` threads.
  spc_smoothed <- graphSmooth(spc, 
                              width = 512, 
                              height = 512, 
                              alpha = 2.0, 
                              neighbors = 8)
  
  return(spc_smoothed)
}

# Dispatch the jobs to the Slurm cluster asynchronously
# future_lapply acts just like lapply, but distributes the work 
# across the compute nodes defined in our `plan()`.
smoothed_results <- future_lapply(spc_list, smooth_wrapper)
```

### What happens under the hood?

1. `future_lapply` breaks the `spc_list` into chunks.
2. It generates batch scripts using `slurm.tmpl` and submits them via `sbatch`.
3. The Slurm scheduler allocates a node with exactly 4 CPU cores for each job.
4. When the R worker starts on the compute node, the `smooth_wrapper` reads the `SLURM_CPUS_PER_TASK` environment variable.
5. It explicitly restricts `RAYON_NUM_THREADS = 4`.
6. Rust's `faer` and `rayon` crates perform the spatial Laplacian smoothing using exactly 4 threads, maximizing performance without violating cluster policies.
7. The results are automatically collected and returned to your master R session.

By combining the inter-node distribution of `future.batchtools` with the intra-node speed of `hySpc.hpc`, you can process massive hyperspectral datasets in a fraction of the time.
