---
title: "Spatial Smoothing with Rust in hySpc.hpc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatial Smoothing with Rust in hySpc.hpc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Hyperspectral imaging data often suffers from spatial noise. A common preprocessing step is to apply a spatial smoother. This vignette demonstrates how to use the `graphSmooth()` function provided by `hySpc.hpc`. 

This implementation leverages Rust's `extendr` framework and an iterative Krylov solver (Conjugate Gradient or BiCGSTAB) to solve the sparse Laplacian equation:

$$
(I + \alpha L) x = b
$$

Where:
* $I$ is the identity matrix.
* $L$ is the graph Laplacian of the spatial grid.
* $\alpha$ is a smoothing parameter.
* $b$ is the noisy input hyperspectral data.
* $x$ is the resulting smoothed data.

## Usage

First, load the required packages.

```{r setup, message=FALSE}
library(hyperSpec)
library(hySpc.hpc)
```

Let's create a simulated noisy hyperspectral image. We'll use a small $10 \times 10$ spatial grid and $5$ wavelength bands for demonstration.

```{r simulate}
set.seed(123)
width <- 10
height <- 10
n_pixels <- width * height
n_bands <- 5

# Create a clean signal (e.g. a simple gradient)
clean_signal <- matrix(rep(1:n_pixels, n_bands), nrow = n_pixels) / n_pixels

# Add Gaussian noise
noisy_signal <- clean_signal + matrix(rnorm(n_pixels * n_bands, sd = 0.2), nrow = n_pixels)

# Wrap it in a hyperSpec object
spc_noisy <- new("hyperSpec", spc = noisy_signal)
```

### Applying the Smoother

To apply the smoother, we simply call `graphSmooth()`, providing the spatial dimensions.

```{r smooth}
# Apply smoothing with alpha = 2.0 and 8-connectivity
spc_smoothed <- graphSmooth(spc_noisy, width, height, alpha = 2.0, neighbors = 8)
```

We can visually compare the noisy and smoothed data.

```{r plot, fig.width=6, fig.height=4}
# Extract the first wavelength band and reshape to a matrix for plotting
noisy_band1 <- matrix(spc_noisy[,,1]@data$spc, nrow = height, ncol = width)
smoothed_band1 <- matrix(spc_smoothed[,,1]@data$spc, nrow = height, ncol = width)

par(mfrow=c(1,2))
image(noisy_band1, main="Noisy Band 1", col=hcl.colors(12, "viridis"))
image(smoothed_band1, main="Smoothed Band 1", col=hcl.colors(12, "viridis"))
```

### Parameters

* **`alpha`**: Controls the amount of smoothing. Higher values result in stronger smoothing.
* **`neighbors`**: The connectivity of the spatial grid. Can be `4` (up, down, left, right) or `8` (includes diagonals).
* **`solver`**: The iterative Krylov solver used by the Rust backend. Can be `"cg"` (Conjugate Gradient, default) or `"bicgstab"` (BiCGSTAB).

```{r strong_smooth}
# Stronger smoothing
spc_strong <- graphSmooth(spc_noisy, width, height, alpha = 10.0, neighbors = 8)
strong_band1 <- matrix(spc_strong[,,1]@data$spc, nrow = height, ncol = width)

par(mfrow=c(1,1))
image(strong_band1, main="Strongly Smoothed (alpha = 10.0)", col=hcl.colors(12, "viridis"))
```

## Performance

By utilizing iterative Krylov solvers natively in Rust, `graphSmooth()` is designed to handle very large images efficiently. Memory overhead is minimized by directly passing memory slices via `extendr`.
